---
title: "Correlations"
author: "Ange Olson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```


```{r, echo = F}
# read in dataset
airbnb <- read.csv("NYC ABS.csv", header = TRUE)

# subset columns for correlation and cast as numeric
airbnb_cor <- airbnb[ , c(10, 11, 12, 14, 15, 16)]
airbnb_cor[ , ] <- lapply(airbnb_cor[ , ], as.numeric)
```

## Correlation Matrix for Airbnb Data
```{r, results = "markup"}
cor(airbnb_cor)
```

No strong correlations with price, but `minimum_nights` and `availability_365`, `number_of_reviews` and `availability_365`, and `calculated_host_listings_count` and `availability_365` show some evidence of positive correlation. 

## Correlation Between Reviews per Month and Total Reviews

```{r, results = "markup"}
plot(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews)
cor.test(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews)
```

# Scatter plot
```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(reviews_per_month, number_of_reviews)) + 
  ggtitle("Number of Reviews VS Reviews per Month") + 
     xlab("Reviews per month") + 
       ylab("Number of reviews") + 
        geom_point(colour = 'blue') + 
          theme_grey()
cor.test(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews, method = "pearson")
```


As expected, correlated since reivews per month is a function of total number of reviews so do not need to look at both.


## Basic Scatter Plots for Correlated Variables

```{r}
plot(x=airbnb_cor$availability_365, y=airbnb_cor$minimum_nights)
plot(x=airbnb_cor$availability_365, y=airbnb_cor$number_of_reviews)
plot(x=airbnb_cor$availability_365, y=airbnb_cor$calculated_host_listings_count)
```

```{r, results = "markup"}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=availability_365, y=minimum_nights)) + ggtitle("Minimum nights vs Availability Scatter plot") + xlab("Availability") + ylab("# Minimum nights") + geom_point(size = 1, shape = 18, color = "black") + ylim(0,400) + stat_cor(method = "pearson", label.x = 275) + geom_smooth(method = lm, se = FALSE, color = "yellow", size = 1.2) + theme_bw()
cor.test(airbnb$availability_365, airbnb$minimum_nights)
```

```{r}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=availability_365, y=number_of_reviews)) + ggtitle("Number of reviews vs Availability") + xlab("Availability") + ylab("# Reviews")  + geom_point(size = 1, shape = 18, color = "black") + ylim(0,800) + stat_cor(method = "pearson", label.x = 275) + geom_smooth(method = lm, se = FALSE, color = "yellow", size = 1.2) + theme_bw()
cor.test(airbnb$availability_365, airbnb$number_of_reviews)
```


```{r}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=availability_365, y=calculated_host_listings_count)) + ggtitle("# Host listings vs Availability") + xlab("Availability") + ylab("Host listings") + geom_point(size = 1, shape = 18, color = "black") + ylim(0,400) + stat_cor(method = "pearson", label.x = 275) + geom_smooth(method = lm, se = FALSE, color = "yellow", size = 1.2) + theme_bw()
cor.test(airbnb$availability_365, airbnb$calculated_host_listings_count)
```

## Correlation Between Number of Reviews (Y) and Price (X) 

```{r, results = "markup"}
plot(y=airbnb$number_of_reviews, x=airbnb$price)
cor.test(y=airbnb$number_of_reviews, x=airbnb$price)
```

```{r, results = "markup"}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=price, y=number_of_reviews,)) + ggtitle("Number of Reviews vs Price Scatter Plot") + xlab("Price ($)") + ylab("Number Of Reviews") + geom_point(size = 1, shape = 18, color = "black") + geom_smooth(method = lm, se = FALSE, color = "yellow", size = 1.2) + theme_bw() + stat_cor(method =  "pearson", label.x = 7500 )
```

No evidence of strong (linear) correlation, but evidence of an inverse relationship between price and reviews (higher price, fewer reviews--possibly because of fewer stays, for which review number is probably a good proxy)

## Histogram Plots for the five variables

# Hist plot for "Reviews per Month" variable

```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(x = reviews_per_month)) + 
  geom_histogram(color="yellow", fill="red") + 
    labs( title = "Histogram for Reviews per Month", 
        x = "Reviews per Month", 
         y = "Frequency" ) +
           theme_dark()
```
# Hist plot for "Number of Reviews" variable

```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(x = number_of_reviews)) + 
  geom_histogram(color="Red", fill="Yellow") + 
    labs( title = "Histogram for Number of Reviews", 
        x = "Number of Reviews", 
         y = "Frequency" ) +
           theme_dark()
```
# Hist plot for "Minimum Nights" variable

```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(x = minimum_nights)) + 
  geom_histogram(color="#E69F00", fill="#56B4E9",) + 
    labs( title = "Histogram for Minimum Nights", 
        x = "Minimum Nights", 
         y = "Frequency" ) +
           theme_dark()
```
# Hist plot for "Calculated Host Listings" variable

```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(x = calculated_host_listings_count)) + 
  geom_histogram(color="white", fill="#E69F00") + 
    labs( title = "Histogram for Count of Calculated Host Listings", 
        x = "Calculated Host Listings", 
         y = "Frequency" ) +
           theme_dark()
```

# Hist plot for "Airbnb Price" variable

```{r, results = 'markup'}
library(ggplot2)
ggplot(airbnb_cor, aes(x = airbnb$price)) + 
  geom_histogram(color="mediumorchid2", fill="yellowgreen") + 
    labs( title = "Histogram for Airbnb Price", 
        x = "Airbnb Price", 
         y = "Frequency" ) +
           theme_dark()
```


#summary of dataset
```{r}
summary(airbnb)
```
#box plot for neighbourhood group vs price
```{r, results='markup'}
library(ggplot2)
ggplot(airbnb, aes(price, factor(neighbourhood_group))) + geom_boxplot(width = 0.7, color = "black", fill = c("light green", "pink","light blue", "yellow", "red")) +labs(title = "Neighbourhood group vs Price Box plot", x = "Price", y = "Neighbourhood")

```
#boxplot for room type vs price
```{r, results='markup'}
library(ggplot2)
ggplot(airbnb, aes(price, factor(room_type))) + geom_boxplot(width = 0.7, color = "black", fill = c("light green", "yellow","light blue")) +labs(title = "Room type vs Price Box plot", x = "Price", y = "Room Type")
```


```{r}

#Load the library
library("ggmap")

#Set your API Key
ggmap::register_google(key = "AIzaSyBuM2zUJqBlgjcki9tYS1emZr3awesSqac")
```

#new york map without need for the API key
```{r}
library(ggplot2)
library(ggmap)
library(tmaptools)
ggmap(get_stamenmap(rbind(as.numeric(paste(geocode_OSM("New York")$bbox))), zoom = 11)) + geom_point(data = airbnb, alpha = .01, aes(x = longitude, y = latitude))
```

#New york map with Lat and Long data by price
```{r, results='markup'}
library(ggplot2)
library(ggmap)
newyork.map <- get_map("New York", zoom = 10, scale = 1, maptype = "terrain", color = "bw")
ggmap(newyork.map) + geom_point(data=airbnb, aes(x=longitude, y=latitude, color = price), size=0.2)
```

#new york map with lat and long data
```{r, results='markup'}
newyork.map <- ggmap(get_map(location= 'Lower Manhattan, New York', 
                       maptype='roadmap', color='bw',source='google',zoom=10)) + 
  geom_point(data=airbnb, aes(x=longitude,y=latitude), color='black',size=0.2)
newyork.map
```


#necessary package for mapping
```{r}
#install.packages("ggmap")
#install.packages("tmaptools")
#install.packages("lubridate")
#install.packages("ggplot2")
#install.packages("data.table")
#install.packages("ggrepel")
#install.packages("dplyr")
#install.packages("data.table")
#install.packages("tidyverse")
#devtools::install_github("dkahle/ggmap", ref = "tidyup", force=TRUE)

```


