---
title: "Midterm EDA"
author: "Group 3: Pavani Samala, Ange Olson, Kowshik Bezawada, Meghana Gantla"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
library(ezids)
```

# Section I: Set-Up and Summary of the Dataset

First, we read in the data and set it up for analysis. The data is mostly cleaned, but we need a subset for calculating correlation, we need to change some data to be categorical, some data to be numerical, and we need to fix the dates so that they aren't read in as characters.

Without doing anything, our dataset is as follows:

```{r, results="markup"}
# read in dataset
airbnb <- data.frame(read.csv("NYC ABS.csv", header = TRUE))

#structure of dataset
str(airbnb)
```


```{r, echo = F}

# edit dates
airbnb$last_review <- as.Date(airbnb$last_review, format = "%m/%d/%Y")

# cast categorical variables
airbnb$neighbourhood_group <- as.factor(airbnb$neighbourhood_group)
airbnb$neighbourhood <- as.factor(airbnb$neighbourhood)
airbnb$room_type <- as.factor(airbnb$room_type)

# cast numerical variables as numeric
airbnb[ , c(10, 11, 12, 14, 15, 16)] <- lapply(airbnb[ , c(10, 11, 12, 14, 15, 16)], as.numeric)

# subset rows for correlation and cast as numeric
airbnb_cor <- airbnb[ , c(10, 11, 12, 14, 15, 16)]
```

After cleaning, our main dataset is described below:

```{r, results='markup'}
str(airbnb)
```

Our secondary dataset (used to measure correlation) is described below:

```{r, results='markup'}
str(airbnb_cor)
```

# Section II: Descriptive Statistics

Our dataset has `r nrow(airbnb)` observations. The continuous variables are described below (NA values omitted):

```{r, results = "markup", echo = F}
xkablesummary(na.omit(airbnb_cor))
```

```{r}
library(DescTools)
```

The standard deviations of the continuous variables are as follows:

*   `price`: `r sd(airbnb_cor$price)`
*   `minimum_nights`: `r sd(airbnb_cor$minimum_nights)`
*   `number_of_reviews`: `r sd(airbnb_cor$number_of_reviews)`
*   `reviews_per_month`: `r sd(na.omit(airbnb_cor$reviews_per_month))`
*   `calcualted_host_listings_count`: `r sd(airbnb_cor$calculated_host_listings_count)`
*   `availability_365`: `r sd(airbnb_cor$availability_365)`

The modes of the continuous variables are as follows:

*   `price`: `r Mode(airbnb_cor$price)`
*   `minimum_nights`: `r Mode(airbnb_cor$minimum_nights)`
*   `number_of_reviews`: `r Mode(airbnb_cor$number_of_reviews)`
*   `reviews_per_month`: `r Mode(na.omit(airbnb_cor$reviews_per_month))`
*   `calcualted_host_listings_count`: `r Mode(airbnb_cor$calculated_host_listings_count)`
*   `availability_365`: `r Mode(airbnb_cor$availability_365)`

The main categorical variables of interest are counted below:

```{r, results='markup'}

#neighborhood group
xkabledply(table(airbnb$neighbourhood_group))

#room_type
xkabledply(table(airbnb$room_type))
```

# Section III: Plots and Graphs

## Histogram for Price (All Observations)

First, we look at the price distribution for the entire dataset:

```{r, results = 'markup'}
library(ggplot2)
ggplot(data=airbnb, aes(price)) + 
  geom_histogram(bins = 100, 
                 col  = "dark blue", 
                 fill = "light blue", 
                alpha = .7) + # opacity
    labs(title = "Histogram of AirBnB Price (All Observations)", 
             x = "AirBnB Price", 
             y = "Frequency") +
     theme_grey()
```


## Q-Q plot for Price (All Observations)
```{r, results='markup'}
qqnorm(airbnb$price, pch = 20, main = "Q-Q Plot for AirBnB Prices (All Observations)")
qqline(airbnb$price, col = "black", lwd = 2)
```

Price is not normally distributed, and there appear to be a large number of outliers. If we remove those outliers, the distribution is close to normal. 

```{r, results='markup'}
airbnb_clean = outlierKD2(airbnb, price, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
```

## Histogram for Price (Outliers Removed)

```{r, results = 'markup'}
library(ggplot2)
ggplot(data = airbnb_clean, aes(price)) +  
  geom_histogram(bins = 100, 
                 col  = "dark blue", 
                 fill = "light blue", 
                alpha = .7) +  # opacity
    labs(title = "Histogram of AirBnB Price (Outliers removed)", 
             x = "AirBnB Price", 
             y = "Frequency") +
     theme_grey()
```

## Q-Q plot for Price (Outliers Removed)
```{r, results='markup'}
qqnorm(airbnb_clean$price, pch = 20, main = "Q-Q Plot for AirBnB Prices (Outliers Removed)")
qqline(airbnb_clean$price, col = "black", lwd = 2)
```


## Scatter Plot for Price and Number of Reviews

Scatter plot for `price` and `number of reviews` without any transformations: 

```{r, results='markup'}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=price, y=number_of_reviews)) + 
  ggtitle("Number of Reviews vs Price Scatter Plot") + 
  xlab("Price ($)") + ylab("Number Of Reviews") + 
  geom_point(size = 1, shape = 18, color = "black") + 
  theme_bw()
```

There appears to be a an exponential relationship (exponential decline) between `price` and `number of reviews`. 


Scatter plot for `price` and `number of reviews` taking the $$\log (reviews)$$ to show linear trend: 

```{r, results='markup'}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=price, y=log(number_of_reviews))) + 
  ggtitle("Number of Reviews vs Price Scatter Plot") + 
  xlab("Price ($)") + ylab("Number Of Reviews (Natural Log)") + 
  geom_point(size = 1, shape = 18, color = "black")
```

## Box Plot for Price and Neighborhood Group

Note: outliers extend past $1,000 per night, graph truncated for visibility and interpretation. 

```{r, results = "markup"}
library(ggplot2)
ggplot(airbnb, aes(price, factor(neighbourhood_group))) + 
  geom_boxplot(color = "black", fill = c("light green", "pink","light blue", "yellow", "orange")) +
  labs(title = "Neighbourhood Group vs Price Box plot", x = "Price", y = "Neighbourhood group") +
  xlim(0, 1000)
```

## Box Plot for Price and Room Type
Note: outliers extend past $1,000 per night, graph truncated for visibility and interpretation. 

```{r, results = "markup"}
library(ggplot2)
ggplot(airbnb, aes(price, factor(room_type))) + 
  geom_boxplot(width = 0.7, color = "black", fill = c("light green", "yellow","light blue")) +
  labs(title = "Room type vs Price Box plot", x = "Price", y = "Room Type") + xlim(0, 1000)
```


## Map of AirBnbs in NY by Neighbourhood group
```{r, results = 'markup'}
#install.packages("plotly")
library(plotly)

fig <- airbnb

fig <- fig %>%

  plot_ly(

    lat = ~latitude,

    lon = ~longitude,

    color = ~neighbourhood_group,
    
    colors = "Set1",

    type = 'scattermapbox')

fig <- fig %>%

  layout(

    mapbox = list(

      style = 'open-street-map',

      zoom =9,

      center = list(lon = -73.97, lat = 40.71))) 


fig
```


## AirBnB Average Price and Number of Listings by Neighborhood:

```{r}
# create subset just for aggregating by mean
airbnb_map <- airbnb[ , c(6, 7, 8, 10)]
airbnb_map_means <- aggregate(.~neighbourhood, airbnb_map, mean)

# create subset for aggregating by count
airbnb_count <- airbnb_map
airbnb_count$count <- 1
airbnb_count <- airbnb_count[, c(1,5)]
airbnb_counter <- aggregate(.~neighbourhood, airbnb_count, sum)

# create full dataset from both subsets
airbnb_map_full <- cbind(airbnb_counter, airbnb_map_means)

# check that union occured correctly, then drop extra neighborhood value
all.equal(airbnb_map_full[, 1], airbnb_map_full[, 3]) # true!
airbnb_map_full <- airbnb_map_full[, -3]
```

```{r, results = 'markup'}
#Load the library
library(ggplot2)
library(ggmap)

#Set your API Key
ggmap::register_google(key = "AIzaSyBuM2zUJqBlgjcki9tYS1emZr3awesSqac")

#map by price 
newyork.map <- get_map("New York", zoom = 10, scale = 1, maptype = "terrain")  
ggmap(newyork.map) +  geom_point(data = airbnb_map_full, aes(x = longitude, y = latitude, colour = price, size = count), alpha = 0.5) + 
  scale_colour_gradientn(colours=rainbow(3)) +
  labs(title = "Average AirBnb Price and Density by Neighborhood", x = "Longitude", y = "Latitude")
```

# Section IV: Correlation and ANOVA Tests

## 4.1: Correlation
### 4.1.1: Correlation Matrix for Airbnb Data

To answer our SMART question, we needed to investigate if any variables other than location or room type seemed to be related to the variable of interest, `price`. The other variables identified are as follows:

* `minimum_nights`
* `number_of_reviews`
* `reviews_per_month`
* `calculated_host_listings_count`
* `availability_365`

As all of these variables are continuous, we have chosen to look at correlation. It is important to note that correlation does not imply causation. However, for the purposes of our SMART question, if for example a variable $x$ is highly and positively correlated with price, we can say that the presence of $x$ is associated with an increase in price and that will be sufficient.

Below we have a correlation matrix and accompanying correlation plot using all complete observations (note: `reviews per month` contains NA values).

```{r, results = "markup"}
loadPkg("faraway")
loadPkg("corrplot")
xkabledply(cor(airbnb_cor, use = "complete.obs"))
airbnb_corplot = cor(airbnb_cor, use = "complete.obs")
corrplot(airbnb_corplot, method = "circle")
```

As shown, no variable of interest shows strong Pearson correlation with price, but `minimum_nights` and `availability_365`, `number_of_reviews` and `availability_365`, and `calculated_host_listings_count` and `availability_365` show some evidence of positive correlation. 

Additionally, as expected, we see strong correlation between the total number of reviews and the reviews per month. This makes sense; we would expect that the more reviews in total a unit has, the more reviews it receives per month. This correlation is significantly significant, as shown in subsection 4.1.2.

While the correlation between `price` and `number_of_reviews` is not high, as shown from our previous scatterplots there does appear to be an exponential relationship between the two variables, with the highest-priced units receiving very few reviews total. In subsection 4.1.3, we see that this relationship is not strong but that there is some statistically significant evidence of correlation. This relationship makes sense; number of reviews can be a proxy for total number of stays in the unit, and one would expect that expensive units are stayed in less frequently. 

### 4.1.2: Correlation Between Reviews per Month and Total Reviews

```{r, results = "markup"}
cor.test(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews)
```


### 4.1.3: Correlation Between Number of Reviews (Y) and Price (X) 

```{r, results = "markup"}
cor.test(y=airbnb$number_of_reviews, x=airbnb$price)
```


## 4.2: ANOVA Tests

### 4.2.1: Testing for Differences in Price by Neighborhood Group

```{r, results = "markup"}
#anova test for price and neighborhood groups
anova_price_group = aov(price ~ neighbourhood_group, data=airbnb)
summary(anova_price_group) -> sum_anova_price_group
xkabledply(sum_anova_price_group, title = "ANOVA result summary for Neighborhood Groups")

tukeyAoV_pg <- TukeyHSD(anova_price_group)
tukeyAoV_pg

```

#### 4.2.1.1 Tukey HSD Results:

At a statistically significant level with $\alpha = 0.05$, we see that there are statistically significant differences between Manhattan and all other boroughs, between Brooklyn and the Bronx, and between Queens and Brooklyn. Differences are not statically significant between Staten Island, Queens, and the Bronx.

This aligns with our previous visual findings, that Manhattan appeared to the most expensive borough with Brooklyn coming in at second. In terms of using location to determine price, we can conclude that certain boroughs are more useful predictors than others. For example, knowing that a unit is in Manhattan as opposed to in the Bronx would be useful, but with no other information, a unit in Staten Island may be priced roughly the same as a unit in Queens. Still, given that at least some locations are priced differently from each other, we can conclude that yes, location does have an effect on price. 


### 4.2.2: Testing for Differences in Price by Room Type

```{r, results = "markup"}
#anova test for price and room type
anova_price_room = aov(price ~ room_type, data=airbnb)
summary(anova_price_room) -> sum_anova_price_room
xkabledply(sum_anova_price_room, title = "ANOVA result summary for Room Type")

tukeyAoV_pr <- TukeyHSD(anova_price_room)
tukeyAoV_pr
```

#### 4.2.2.1 Tukey HSD Results:

At a statistically significant level with $\alpha = 0.05$, we see that there are statistically significant differences across all room types. This lines up with our visual representations as well, where we saw that an entire home or apartment was significantly preferred (priced higher) than either room. While not as visually obvious, a private room is also priced statistically significantly higher than a shared room. We can conclude that room type does likely have an effect on price.




