---
title: "Correlations"
author: "Group 3: Pavani Samala, Ange Olson, Kowshik Bezawada, Meghana Gantla"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
library(ezids)
```

# Section I: Set-Up and Summary of the Dataset

First, we read in the data and set it up for analysis. The data is mostly cleaned, but we need a subset for calculating correlation, we need to change some data to be categorical, some data to be numerical, and we need to fix the dates so that they aren't read in as characters.

Without doing anything, our dataset is as follows:

```{r, echo = F}

```


```{r, echo = F}
# read in dataset
airbnb <- data.frame(read.csv("NYC ABS.csv", header = TRUE))

# edit dates
airbnb$last_review <- as.Date(airbnb$last_review, format = "%m/%d/%Y")

# cast categorical variables
airbnb$neighbourhood_group <- as.factor(airbnb$neighbourhood_group)
airbnb$neighbourhood <- as.factor(airbnb$neighbourhood)
airbnb$room_type <- as.factor(airbnb$room_type)

# cast numerical variables as numeric
airbnb[ , c(10, 11, 12, 14, 15, 16)] <- lapply(airbnb[ , c(10, 11, 12, 14, 15, 16)], as.numeric)

# subset rows for correlation and cast as numeric
airbnb_cor <- airbnb[ , c(10, 11, 12, 14, 15, 16)]
```

After cleaning, our main dataset is described below:

```{r, echo = F}

```

Our secondary dataset (used to measure correlation) is described below:

```{r, echo = F}

```

# Section II: Descriptive Statistics


# Section III: Plots and Graphs

## Scatter Plot for Price and Number of Reviews

Scatter plot for `price` and `number of reviews`: 
```{r, results = "markup"}
library(ggplot2)
library(ggpubr)
ggplot(airbnb, aes(x=price, y=number_of_reviews,)) + ggtitle("Number of Reviews vs Price Scatter Plot") + xlab("Price ($)") + ylab("Number Of Reviews") + geom_point(size = 1, shape = 18, color = "black") + geom_smooth(method = lm, se = FALSE, color = "yellow", size = 1.2) + theme_bw() + stat_cor(method =  "pearson", label.x = 7500 )
```

## Box Plot for Price and Neighborhood Group

```{r, results = "markup"}
library(ggplot2)
ggplot(airbnb, aes(price, factor(neighbourhood_group))) + geom_boxplot(width = 0.7, color = "black", fill = c("light green", "pink","light blue", "yellow", "red")) +labs(title = "Neighbourhood group vs Price Box plot", x = "Price", y = "Neighbourhood")
```

## Box Plot for Price and Room Type

```{r, results = "markup"}
library(ggplot2)
ggplot(airbnb, aes(price, factor(room_type))) + geom_boxplot(width = 0.7, color = "black", fill = c("light green", "yellow","light blue")) +labs(title = "Room type vs Price Box plot", x = "Price", y = "Room Type")
```

## Map of NY by Price

### Option 1: 
```{r, results='markup'}
library(ggmap)
library(tmaptools)
ggmap(get_stamenmap(rbind(as.numeric(paste(geocode_OSM("New York")$bbox))), zoom = 10)) + geom_point(data = airbnb, aes(x = longitude, y = latitude))
```

# Section IV: Correlation and ANOVA Tests

## Correlation
### Correlation Matrix for Airbnb Data
```{r, results = "markup"}
loadPkg("faraway")
loadPkg("corrplot")
xkabledply(cor(airbnb_cor))
airbnb_corplot = cor(airbnb_cor, use = "complete.obs")
corrplot(airbnb_corplot, method = "circle")
```

No strong correlations with price, but `minimum_nights` and `availability_365`, `number_of_reviews` and `availability_365`, and `calculated_host_listings_count` and `availability_365` show some evidence of positive correlation. 

### Correlation Between Reviews per Month and Total Reviews

```{r, results = "markup"}
plot(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews)
cor.test(x=airbnb_cor$reviews_per_month, y=airbnb_cor$number_of_reviews)
```

As expected, correlated since reviews per month is a function of total number of reviews so do not need to look at both.


### Correlation Between Number of Reviews (Y) and Price (X) 

```{r, results = "markup"}
plot(y=airbnb$number_of_reviews, x=airbnb$price)
cor.test(y=airbnb$number_of_reviews, x=airbnb$price)
```

No evidence of strong (linear) correlation, but evidence of an inverse relationship between price and reviews (higher price, fewer reviews--possibly because of fewer stays, for which review number is probably a good proxy)


## ANOVA Tests

### Testing for Differences in Price by Neighborhood Group

```{r aov, results='markup'}
#anova test for price and neighborhood groups
anova_price_group = aov(price ~ neighbourhood_group, data=airbnb)
anova_price_group
summary(anova_price_group) -> sum_anova_price_group
xkabledply(sum_anova_price_group, title = "ANOVA result summary for Neighborhood Groups")

tukeyAoV_pg <- TukeyHSD(anova_price_group)
tukeyAoV_pg

```


### Testing for Differences in Price by Room Type

```{r aov, results='markup'}
#anova test for price and room type
anova_price_room = aov(price ~ room_type, data=airbnb)
anova_price_room
summary(anova_price_room) -> sum_anova_price_room
xkabledply(sum_anova_price_room, title = "ANOVA result summary for Room Type")

tukeyAoV_pr <- TukeyHSD(anova_price_room)
tukeyAoV_pr
```





